Create data sets for modeling

### 2016-06-29 portfolio/analysis/4a-createData_linear_modeling.Rmd
Merge annual and differenced data creation scripts

---
title: "createData_linear_modeling"
author: ""
date: "June 05, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

@@

## Load cleaned data
```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
library(date)

#load("cfecCleaned2.Rdata")

@@

if (Sys.info()[["user"]] != "seananderson") {
  region<-read.csv("areas/regions.csv")
  region<-region[,c(1:2)] #don’t include “assigned by” column
  cfec<-left_join(cfec,region)
#cfec$region <- replace(cfec$region, is.na(cfec$region), "Unassigned")
}
```

@@

https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0

@@

### Function definiton
We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHI) to calculate the entropy-based inverse, which is a proxy for effective species diversity.

@@

Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.

@@

## Filtering out people with 10+ years of fishing. Drops rows from 13.7 to 10 million.
```{r}
cfecnew = group_by(cfecnew, p_holder) %>%
  mutate(nyr = length(unique(year))) %>%
  filter(nyr >= 10)
```

@@

###For variability, any person in a year will need to have 10% or more of their revenue to come from each permit (otherwise doesn't count as a strategy)

@@

CUTOFF = 50
write.csv(rev(sort(table(topstrategies$strategy))), file = "../data-generated/strategies.csv")
# cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
```

@@

# fixme I assume we no longer want to do this? -SA
#cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
#  filter(npermit == 1)

@@

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
#cfecAnnual$p2.prev = cfecAnnual$p2[idx]

cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev
# idx = which(cfecAnnual$strategy != cfecAnnual$strategy.prev)

# Filter out salmon people
cfecAnnual$salmon = ifelse(substr(cfecAnnual$permit,1,1)=="S", TRUE, FALSE)

# cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
# cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]

#cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
#cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]

# saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
# saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")
cfecAnnual <- as_data_frame(cfecAnnual)
nrow(cfecAnnual)
saveRDS(cfecAnnual, file="../data-generated/cfec-for-modeling.rds")

>>>>

---
title: "Create data sets for modeling"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

## Differenced vs. undifferenced datasets:

1. Excluding observations that we don't have consecutive years for

@@

## Load cleaned data

```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
library(date)

@@

if (Sys.info()[["user"]] != "seananderson") {
  region<-read.csv("areas/regions.csv")
  region<-region[,c(1:2)] #don’t include “assigned by” column
  cfec<-left_join(cfec,region)
}
```

@@

<https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0>

@@

## Function definition

We need to define a few functions for calculating diversity. We'll use the
Simpson's diversity index (HHI) to calculate the entropy-based inverse, which
is a proxy for effective species diversity.

@@

Now we'll filter out (i.e. keep) individuals who only used 1 boat in a year, and boats with
only one permit holder fishing in a year.

@@

## Combine some strategies

@@

Any person in a year will need to have 10% or more of their revenue to come
from each permit (otherwise doesn't count as a strategy).

@@

write.csv(rev(sort(table(topstrategies$strategy))), file = "../data-generated/strategies.csv")
```

@@

@@

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
cfecAnnual$weight.prev = cfecAnnual$weight[idx]
cfecAnnual$days.prev = cfecAnnual$days[idx]

# idx = which(cfecAnnual$strategy != cfecAnnual$strategy.prev)
```

Add some identifying columns:

```{r}
cfecAnnual$area = substr(cfecAnnual$p_fshy,4,4)
cfecAnnual$area[cfecAnnual$area %in% c("A","C")] = "Southeast"
cfecAnnual$area[cfecAnnual$area == "B"] = "Statewide"
cfecAnnual$area[cfecAnnual$area == "D"] = "Yakutat"
cfecAnnual$area[cfecAnnual$area == "E"] = "PWS"
cfecAnnual$area[cfecAnnual$area == "H"] = "CookInlet"
cfecAnnual$area[cfecAnnual$area == "J"] = "Westward"
cfecAnnual$area[cfecAnnual$area == "K"] = "Kodiak"
cfecAnnual$area[cfecAnnual$area == "L"] = "Chignik"
cfecAnnual$area[cfecAnnual$area == "M"] = "AlaskaPeninsula"
cfecAnnual$area[cfecAnnual$area == "N"] = "Nelson"
cfecAnnual$area[cfecAnnual$area == "Q"] = "BeringSea"
cfecAnnual$area[cfecAnnual$area == "T"] = "BristolBay"
cfecAnnual$area[cfecAnnual$area == "X"] = "Kotzebue"
cfecAnnual$area[cfecAnnual$area == "Z"] = "NortonSound"

cfecAnnual$spec = substr(cfecAnnual$p_fshy,1,1)
cfecAnnual$spec[cfecAnnual$spec == "B"] = "Halibut"
cfecAnnual$spec[cfecAnnual$spec == "C"] = "Sablefish"
cfecAnnual$spec[cfecAnnual$spec %in% c("D","E","K","T")] = "Crab"
cfecAnnual$spec[cfecAnnual$spec %in% c("F")] = "Freshwater"
cfecAnnual$spec[cfecAnnual$spec %in% c("G","H","L")] = "Herring"
cfecAnnual$spec[cfecAnnual$spec %in% c("I")] = "Lingcod"
cfecAnnual$spec[cfecAnnual$spec %in% c("A","J","N","R","W")] = "Shellfish"
cfecAnnual$spec[cfecAnnual$spec %in% c("O","P","Q","U","Z")] = "Invert"
cfecAnnual$spec[cfecAnnual$spec %in% c("M")] = "Misc"
cfecAnnual$spec[cfecAnnual$spec %in% c("S")] = "Salmon"
cfecAnnual$area_spec = paste(cfecAnnual$area, cfecAnnual$spec, sep="_")
```

Save the data for modeling:

```{r}
# make sure the data are tidy:
cfecAnnual <- as_data_frame(cfecAnnual)

## Filtering out people with 10+ years of fishing. Drops rows from 13.7 to 10 million.
## (now doing this at the beginning of the modeling scripts)
# cfecAnnual = group_by(cfecAnnual, p_holder) %>%
#   mutate(nyr = length(unique(year))) %>%
#   filter(nyr >= 10)

saveRDS(cfecAnnual, file="../data-generated/cfec-annual-for-modeling.rds")

cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev
cfecAnnual.diff = cfecAnnual[is.na(cfecAnnual$specDiv.change)==FALSE, ]
saveRDS(cfecAnnual.diff, file="../data-generated/cfec-diff-for-modeling.rds")
```
